Mangrove forests along tropical coastlines frequently suffer severe damage due to storms. The mangrove trees are extremely sensitive to environmental stressors such as water salinity and tidal inundation. We propose an agent-based model for the prediction of the regenerative behavior of mangrove stands consisting of the native species and the planted or non-native species in a fragmented habitat, with the use of spatio-temporal colored noise for stochastic seedling dispersal. This study aims to accurately model the growth of mangrove populations while experiencing said factors and subject to storm damage. It uses Salmo and Juanico's agent-based model for mangrove growth, along with García-Ojalvo, Sancho, and Ramírez-Piscina's algorithm for spatiotemperal colored noise. Stochastic experiments were carried out in a shoreline habitat with an existing native population (\textit{Avicennia} and  textit{Sonneratia}) of varying ages and a larger population of planted seedlings  (\textit{Rhizophora}). 

A sensitivity analysis shows that in a run lasting 5000 days, there are two outcomes depending on what variable is varying in the growth equation. When either of the allometric parameters $\alpha$ or $\beta$ (related to tree height and crown radius) are varying, the native species dominates the environment, with both species taking roughly 3.5 years for trees to recover from storms. When everything is constant, or $\lambda$ (allometric parameter related to maximum height) is varying, the non-native species dominates the environment. 
 
In a Monte Carlo experiment wherein the allometric parameters, correlation time and diffusion rate (both used in the spatio-temporal coloured noise associated with seedling dispersal chance) are varying, the native species dominates nearly 77\% of the time with an average recovery time of 3.98 years.